---
title: "data_exploration"
author: "Susan Campbell"
date: "2024-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data descriptions

### Exposure data

These data were provided by Dr. Yukiko Hirabayashi's team. Key features:

-   Temporal coverage and resolution: 2071-2100 with resolution 0.25 \* 0.25 degrees (

-   Spatial coverage and resolution: globe with high resolution (point data)

-   Measure of flood frequency: Return period (number of years) for 10-, 30- and 100-year flood events (according to historical 1971-2000 data)

-   Measure of flood intensity: none

-   Future scenarios: SSP126 ("sustainability" at 1.0-1.8 C warming by 2100), SSP 245 ("middle of road" at 1.3-2.4 C by 2100), and SSP 585 ("fossil fuels" at 3.3-5.7 C by 2100)

For more information about its origins, see: Hirabayashi, Y., Tanoue, M., Sasaki, O., Zhou, X., & Yamazaki, D. (2021). Global exposure to flooding from the new CMIP6 climate model projections. *Scientific Reports*, *11*(1), 3740.

### Hazard data

Describe!

### Country and region data

Describe!

### Population data

Describe!

## Load libraries and set directory

```{r libs, echo=FALSE}

library(tidyverse)
library(tiff)
library(raster)
library(rasterVis)
library(readr)
library(sf)
library(terra)
library(rmapshaper)
library(viridis)
library(cowplot)
library(RColorBrewer)

# specify directories
dir <- "/Users/SusanCampbell/UW_HMS_R/capstone/"
setwd(dir) 

# load functions in my process_input_functions.R
source(paste0(dir, "/processing_functions.R")) 

```

## Load data and set params

Set values related to return period and time window

```{r set_time, echo=FALSE}

# set values related to scenario
initial_return_period <- 30
climate_future <- "SSP585"
flood_size <- paste0(initial_return_period, "yr")

# set seed for reproducibility
set.seed(48)

# set values related to time of interest
time_window <- 100
start_year <- 2024
projection_year <- 2085 # midway between 2071 and 2100
end_year <- start_year + time_window

# locations to save results
grid_results_dir <- paste0(dir, "results/", flood_size, climate_future, "/grid")
country_results_dir <- paste0(dir, "results/", flood_size, climate_future, "/country")

```

Read in other datasets

```{r load_data, echo=FALSE}

# load exposures data: Hirabayashi et al. 2021 ------------------------

# load data for SSP 126 scenario
# rp_10yr_ssp126 <- rast(paste0(dir, "data/ssp126/RP_for_10yr_flood_storage_ssp126.tif"))
# rp_30yr_ssp126 <- rast(paste0(dir, "data/ssp126/RP_for_30yr_flood_storage_ssp126.tif"))
# rp_100yr_ssp126 <- rast(paste0(dir, "data/ssp126/RP_for_100yr_flood_storage_ssp126.tif"))
# 
# # load data for SSP 245 scenario
# rp_10yr_ssp245 <- rast(paste0(dir, "data/ssp245/RP_for_10yr_flood_storage_ssp245.tif"))
# rp_30yr_ssp245 <- rast(paste0(dir, "data/ssp245/RP_for_30yr_flood_storage_ssp245.tif"))
# rp_100yr_ssp245 <- rast(paste0(dir, "data/ssp245/RP_for_100yr_flood_storage_ssp245.tif"))
# 
# # load data for SSP 585 scenario
# rp_10yr_ssp585 <- rast(paste0(dir, "data/ssp585/RP_for_10yr_flood_storage_ssp585.tif"))
# rp_30yr_ssp585 <- rast(paste0(dir, "data/ssp585/RP_for_30yr_flood_storage_ssp585.tif"))
# rp_100yr_ssp585 <- rast(paste0(dir, "data/ssp585/RP_for_100yr_flood_storage_ssp585.tif"))


# load country shapefiles and region hierarchies ---------------------

# from IHME
country_boundaries <- st_read("data/GBD_countries/lbd_standard_admin_0_simplified.shp")
names(country_boundaries) <- c("code", "name", "loc_id", "geo_id", "geometry")

country_id_names_dict <- country_boundaries %>%
  as.data.frame() %>%
  dplyr::select("code", "name", "loc_id") %>%
  mutate(index = row_number())

# GBD location hierarchies: Global burden of disease (GBD) data and Tools Guide: Full list of locations, diseases, injuries, and risk factors. Institute for Health Metrics and Evaluation. (2024). https://www.healthdata.org/research-analysis/about-gbd/gbd-data-and-tools-guide
location_hierarchy <- read_csv("data/gbd_location_hierarchy.csv")
country_region_sdi_dict <- read.csv(file = "data/location_hierarchy.csv")


# load in population data --------------------------------------------

# from WorldPop, processed and corrected by IHME
# load population data from with 0.0417 * 0.0417 degree resolution (~5km at equator)
pop <- rast(paste0(dir, "data/worldpop_total_1y_2024_00_00.tif"))

region_order <- c("Eastern Europe", "Central Europe", "Southern Sub-Saharan Africa", 
                  "North Africa and Middle East", "Central Asia", NA, "Western Europe", 
                  "Australasia", "Southern Latin America", "High-income North America", 
                  "Western Sub-Saharan Africa", "Caribbean", "East Asia", "Central Latin America", 
                  "Tropical Latin America", "Eastern Sub-Saharan Africa", "Andean Latin America", 
                  "Oceania", "South Asia", "High-income Asia Pacific", "Central Sub-Saharan Africa", 
                  "Southeast Asia")

```

Specify whether to write intermediate results to tif files, whether to generate figures, etc.

```{r read_and_write_settings, echo=FALSE}

# if TRUE, grid-level results will be read in from tif files instead of generated anew
read_grid_results <- TRUE
# if TRUE, grid-level results will be saved as tif files
write_grid_results <- FALSE
# if TRUE, figures (maps and histograms) of grid-level results will be saved as pdf files
generate_grid_figures <- TRUE

```

## Grid results

### Generate simulated flood occurrence

From return period, generate simulated binary indicators of flood occurrence for each of the next 100 years. Memory is an issue here. That's why we're testing with just a random sample (0.1%!) from the input raster.

```{r flood_simulation, echo=FALSE}

# generate time series with changing return period ---------------------------

# note: in a test of Option 1, it took ~38 minutes to generate 100 years 
  # for 100% of raster for one of the nine magnitude/climate scenarios

if(read_grid_results == TRUE) {
  
  flood_occurrence_stack <- rast(paste0(
    grid_results_dir, "/output_", flood_size, "_", climate_future,
    "_flood_occurrence.tif"))
  
} else {
  
  # apply function to all cells in raster
  # see process_inputs_functions.R
  rp_slope_raster <- app(rp_100yr_ssp245, 
                         function(x) calculate_rp_slope(projected_rp = x))
  
  # initialize list to store yearly flood draw rasters
  yearly_flood_rasters <- vector("list", time_window)
  
  # apply binomial flood draws to all cells for a given year
  for (year in start_year:(start_year + time_window - 1)) {
    year_index <- year - start_year + 1
    print(paste("Processing year:", year))
    
    yearly_flood_rasters[[year_index]] <- suppressWarnings(
      app(rp_slope_raster, function(cell_rp_slope) {
        sapply(cell_rp_slope, function(x) {
          calculate_year_specific_rp(calculation_year = year, cell_rp_slope = x)
        })
      }))
  }
  
  # convert the list of rasters to a SpatRaster
  flood_occurrence_stack <- rast(yearly_flood_rasters)
  
}


# save to file -------------------------------------------------------------------

if (write_grid_results == TRUE) {
  writeRaster(flood_occurrence_stack,
              paste0(grid_results_dir, "/output_", flood_size, "_", climate_future,
                     "_flood_occurrence.tif"),
              filetype = "GTiff", overwrite = TRUE)
}


# check for plausibility: sum flood occurrences per cell ---------------------------------

# choose year(s) to plot as an example
early_year <- 3
late_year <- 97

map_floods_early <- plot_number_floods_map(flood_occurrence_stack, year = early_year, 
                        title = paste("Cells with", flood_size,
                                      "floods in", climate_future, "(year", early_year, ")"))

map_floods_late <- plot_number_floods_map(flood_occurrence_stack, year = late_year, 
                        title = paste("Cells with", flood_size,
                                      "floods in", climate_future, "(year", late_year, ")"))


hist_floods_early <- plot_number_floods_hist(flood_occurrence_stack, year = early_year, 
                         title = paste("Cells with", flood_size,
                                      "floods in", climate_future, "(year", early_year, ")"))

hist_floods_late <- plot_number_floods_hist(flood_occurrence_stack, year = late_year, 
                         title = paste("Cells with", flood_size,
                                      "floods in", climate_future, "(year", late_year, ")"))


if(generate_grid_figures == TRUE) {
  
  ggsave(paste0(grid_results_dir,
                "/map_floods_year", early_year, "_", flood_size, climate_future, ".pdf"),
    map_floods_early, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(grid_results_dir,
                "/map_floods_year", late_year, "_", flood_size, climate_future, ".pdf"),
    map_floods_late, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(grid_results_dir,
                "/hist_floods_year", early_year, "_", flood_size, climate_future, ".pdf"),
    hist_floods_early, 
    width = 6,
    height = 4,
    dpi = 300)
  
  ggsave(paste0(grid_results_dir,
                "/hist_floods_year", late_year, "_", flood_size, climate_future, ".pdf"),
    hist_floods_late, 
    width = 6,
    height = 4,
    dpi = 300)
  
}


# plot map aggregated over time ----------------------------------------------------------
# see process_inputs_functions.R
flood_sum_raster <- sum_over_time(flood_occurrence_stack)

flood_occurrence_over_time <- plot_occurrence_over_time_map(flood_sum_raster,
                                                  title = paste("Sum of", 
                                                                flood_size,"floods, over next",
                                                                time_window, "years, in",
                                                                climate_future),
                                                  fixed_binned_scales = FALSE)

flood_occurrence_over_time_comparable <- plot_occurrence_over_time_map(flood_sum_raster,
                                                  title = paste("Sum of", 
                                                                flood_size,"floods, over next",
                                                                time_window, "years, in",
                                                                climate_future),
                                                  fixed_binned_scales = TRUE)

if(generate_grid_figures == TRUE) {
  
  ggsave(paste0(grid_results_dir,
                "/map_floods_100_years", "_", flood_size, climate_future, ".pdf"),
    flood_occurrence_over_time, 
    width = 9,
    height = 4,
    dpi = 1000)
  
    ggsave(paste0(grid_results_dir,
                "/map_floods_100_years", "_", flood_size, climate_future, "_comparable.pdf"),
    flood_occurrence_over_time_comparable, 
    width = 9,
    height = 4,
    dpi = 1000)
}

```

### Process population density data

Aggregate population to level of spatial resolution of flood grid

```{r population_to_coarse_grid, echo=FALSE}

# calculate difference in resolution between flood grid and population grid
res_factor <- round(res(flood_occurrence_stack) / res(pop))

# ensure res_factor is the same for both x and y directions
if (res_factor[1] != res_factor[2]) {
  stop("Resolution difference between flood and population grids is not uniform in x and y directions.")
}

# sum population data to level of resolution of flood grid
pop_aggregated <- aggregate(pop, fact = res_factor[1], 
                            fun = sum, expand = TRUE)

pop_country_aggregated <- project(pop_aggregated, crs(country_boundaries))

# aaggregate population to country level
pop_country_aggregated <- terra::extract(pop_country_aggregated, 
                                         country_boundaries[, c("name", "loc_id", "geometry")],
                                         fun = sum, na.rm = TRUE, method = 'simple')
names(pop_country_aggregated) <- c("ID", "worldpop")


# check that resolutions now match
if (!all.equal(res(flood_occurrence_stack), res(pop_aggregated))) {
  stop("Aggregation did not match flood and population resolutions correctly.")
}

# # double check that this aggregates as expected by spot checking a few cells
# # open old raster and new raster, choose coordinates, and make sure they are a sum
# # google a few coordinates for a rough triple check
# get_pop <- function(lat, long, pop_type) {
#   place_coords <- cbind(long, lat)  # Note that longitude is first, then latitude
#   place_point <- vect(place_coords, crs = crs(pop_type))
#   place_pop_value <- extract(pop_type, place_point)
#   return(place_pop_value)
# }
# 
# # square area of Beijing: 16,808 km2. population: 21,540,000 (2018)
# beijing_agg <- get_pop(39.9042, 116.4074, pop_aggregated)
# beijing <- get_pop(39.9042, 116.4074, pop)
# 
# # square area of Bozeman MT: 49.6 km2. population: 56,123 (2022)
# bozemanMT_agg <- get_pop(45.6793, -111.0373, pop_aggregated)
# bozemanMT <- get_pop(45.6793, -111.0373, pop)


# plotting checks -------------------------------------------------------------------

map_pop <- plot_pop_map(pop_aggregated)
hist_pop <- plot_pop_hist(pop_aggregated)

if(generate_grid_figures == TRUE) {
  
  ggsave(paste0(grid_results_dir,
                "/map_pop_", flood_size, climate_future, ".pdf"),
    map_pop, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(grid_results_dir,
                "/hist_pop_", flood_size, climate_future, ".pdf"),
    hist_pop, 
    width = 6,
    height = 4,
    dpi = 300)
  
}

```

### Calculate people exposed

Calculate number of people exposed to each flood-year per cell

```{r exposed_pop_in_grid, echo=FALSE}

if(read_grid_results == TRUE) {
  
  # read in from file if needed
  exposed_stack <- rast(paste0(
    grid_results_dir, "/output_", flood_size, "_", climate_future,
    "_exposures.tif"))
  
} else {

  # calculate exposed population for each year (multiply flood occurrences by population)
  exposed_list <- vector("list", nlyr(flood_occurrence_stack))
  
  # pop_aggregated should be multiplied onto every cell within each layer of   
  # flood_occurrence_stack
  for (i in 1:nlyr(flood_occurrence_stack)) {
    exposed_list[[i]] <- flood_occurrence_stack[[i]] * pop_aggregated
  }
  
  exposed_stack <- rast(exposed_list)

}

# write file to tif -------------------------------------------------------------------

# write file to .tif if needed
if (write_grid_results == TRUE) {
  writeRaster(exposed_stack,
            paste0(grid_results_dir, "/output_", flood_size, "_", climate_future,
                  "_exposures.tif"),
            filetype = "GTiff", overwrite = TRUE)
}


# example year maps and histograms ---------------------------------------------------------

# choose year(s) to plot as an example
early_year <- 3
late_year <- 97

map_exposed_early <- plot_number_exposed_map(exposed_stack, year = early_year, 
                        title = paste("People exposed to", flood_size,
                                      "floods in", climate_future, "(year", early_year, ")"),
                        fixed_binned_scales = FALSE)
map_exposed_early_comparable <- plot_number_exposed_map(exposed_stack, year = early_year, 
                        title = paste("People exposed to", flood_size,
                                      "floods in", climate_future, "(year", early_year, ")"),
                        fixed_binned_scales = TRUE)

map_exposed_late <- plot_number_exposed_map(exposed_stack, year = late_year, 
                        title = paste("People exposed to", flood_size,
                                      "floods in", climate_future, "(year", late_year, ")"),
                        fixed_binned_scales = FALSE)
map_exposed_late_comparable <- plot_number_exposed_map(exposed_stack, year = late_year, 
                                title = paste("People exposed to", flood_size,
                                          "floods in", climate_future, "(year", late_year, ")"),
                                fixed_binned_scales = TRUE)


hist_exposed_early <- plot_number_exposed_hist(exposed_stack, year = early_year, 
                         title = paste("People exposed to", flood_size,
                                      "floods in", climate_future, "(year", early_year, ")"))

hist_exposed_late <- plot_number_exposed_hist(exposed_stack, year = late_year, 
                         title = paste("People exposed to", flood_size,
                                      "floods in", climate_future, "(year", late_year, ")"))


if(generate_grid_figures == TRUE) {
  
  ggsave(paste0(grid_results_dir,
                "/map_exposed_year", early_year, "_", flood_size, climate_future, ".pdf"),
    map_exposed_early, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(grid_results_dir,
                "/map_exposed_year", late_year, "_", flood_size, climate_future, ".pdf"),
    map_exposed_late, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(grid_results_dir,
                "/hist_exposed_year", early_year, "_", flood_size, climate_future, ".pdf"),
    hist_exposed_early, 
    width = 6,
    height = 4,
    dpi = 300)
  
  ggsave(paste0(grid_results_dir,
                "/hist_exposed_year", late_year, "_", flood_size, climate_future, ".pdf"),
    hist_exposed_late, 
    width = 6,
    height = 4,
    dpi = 300)
  
}


# aggregated over time maps -------------------------------------------------------------

# see process_inputs_functions.R
exposure_sum_raster <- sum_over_time(exposed_stack)

exposure_over_time <- plot_exposure_over_time_map(exposure_sum_raster,
                                                  title = paste("Sum of people exposed to", 
                                                                flood_size,"floods, over next",
                                                                time_window, "years, in",
                                                                climate_future),
                                                  fixed_binned_scales = FALSE)

exposure_over_time_comparable <- plot_exposure_over_time_map(exposure_sum_raster,
                                                  title = paste("Sum of people exposed to", 
                                                                flood_size,"floods, over next",
                                                                time_window, "years, in",
                                                                climate_future),
                                                  fixed_binned_scales = TRUE)

if(generate_grid_figures == TRUE) {
  
  ggsave(paste0(grid_results_dir,
                "/map_exposed_100_years", "_", flood_size, climate_future, ".pdf"),
    exposure_over_time, 
    width = 9,
    height = 4,
    dpi = 1000)
  
    ggsave(paste0(grid_results_dir,
                "/map_exposed_100_years", "_", flood_size, climate_future, "_comparable.pdf"),
    exposure_over_time_comparable, 
    width = 9,
    height = 4,
    dpi = 1000)
}

```

### Calculate fraction of people exposed

```{r fraction_exposed, echo=FALSE}

if(read_grid_results == TRUE) {
  
  fraction_country_exposed <- read.csv(paste0(country_results_dir, 
                                              "/fraction_country_exposed_",
                                              flood_size, climate_future, ".csv"))
} else {
  
  # calculate proportion of each country exposed to floods for each year
  fraction_country_exposed <- aggregate_exposure_to_country(exposed_stack, 
                                                            country_boundaries,
                                                            pop_country_aggregated)
}

if(write_grid_results == TRUE) {
  
  write.csv(fraction_country_exposed, 
            file = paste0(country_results_dir, "/fraction_country_exposed_",
                          flood_size, climate_future, ".csv"),
            row.names = FALSE)
  
}

```

## Analyses and visualizations

```{r scenario_summary_plots, echo=FALSE}

# stitch together maps ---------------------------------------------------------------------

# create a list of the plots to include
plots <- list(map_floods_early, map_floods_late, flood_occurrence_over_time, 
              map_exposed_early, map_exposed_late, exposure_over_time, 
              map_pop)

scenario_summary_plot <- plot_grid(plotlist = plots, ncol = 3, nrow = 3)

# save as pdf
save_plot(filename = paste0(grid_results_dir, "/scenario_summary_", 
                            flood_size, "_", climate_future,".pdf"), 
          plot = scenario_summary_plot, 
          dpi = 2400,
          base_height = 18, base_width = 30)

```

## Flood frequency changes

```{r region_summaries, echo=FALSE}

# calculate flood counts for the first 50 years and the second 50 years
flood_counts_first_half_century <- count_floods(flood_occurrence_stack, 1, 50)
flood_counts_second_half_century <- count_floods(flood_occurrence_stack, 51, time_window)

# extract and summarize flood counts for communities within each country for the first 50 years
country_flood_counts_first_half_century <- extract_flood_counts(flood_counts_first_half_century,
                                                                country_boundaries)
summary_floods_first_half_century <- summarize_flood_counts(country_flood_counts_first_half_century,
                                                            country_boundaries,
                                                            country_region_sdi_dict)

# extract and summarize flood counts for communities within each country for the second 50 years
country_flood_counts_second_half_century <- extract_flood_counts(flood_counts_second_half_century,
                                                                 country_boundaries)
summary_floods_second_half_century <- summarize_flood_counts(country_flood_counts_second_half_century,
                                                             country_boundaries,
                                                             country_region_sdi_dict)

# combine results for both time periods
summary_floods_first_half_century$time_period <- "First 50 Years"
summary_floods_second_half_century$time_period <- "Second 50 Years"

tabulated_floods <- bind_rows(summary_floods_first_half_century, 
                              summary_floods_second_half_century) %>%
  pivot_longer(cols = -c(region_name, time_period),
               names_to = "community_floods_tally",
               values_to = "proportion")
tabulated_floods$community_floods_tally <- factor(tabulated_floods$community_floods_tally, 
                                                  levels = c("0", "1-2", "3-4", "5-7", 
                                                             "8-10", "11-15", ">15"))
tabulated_floods$region_name <- factor(tabulated_floods$region_name, 
                                       levels = region_order)

tabulated_floods$community_floods_tally <- factor(tabulated_floods$community_floods_tally,
                                                  levels = c("0", "1-2", "3-4", "5-7", 
                                                             "8-10", "11-15", ">15"))



num_colors <- length(unique(tabulated_floods$community_floods_tally))

flood_first_half_stackedbar <- ggplot(data = filter(tabulated_floods, 
                                        time_period == "First 50 Years" & !is.na(region_name)),
                                      aes(x = region_name, y = proportion, 
                                          fill = community_floods_tally)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues", name = "Number of \nfloods") +
  labs(title = "Flood frequency in communities of each region, first half of next century",
       x = NULL,
       y = NULL) +
  coord_flip() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(-1, 1, by = 0.2)) +
  theme(axis.text = element_text(size = 14),
        legend.position = "none")
ggsave(paste0(country_results_dir,
                "/flood_freq_first_half_century", "_", flood_size, climate_future, ".pdf"),
    flood_first_half_stackedbar, 
    width = 6,
    height = 7,
    dpi = 1000)

flood_second_half_stackedbar <- ggplot(data = filter(tabulated_floods, 
                                        time_period == "Second 50 Years" & !is.na(region_name)), 
                                       aes(x = region_name, y = proportion, 
                                           fill = community_floods_tally)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues", name = "Number of \nfloods") +
  labs(title = "Flood frequency in communities of each region, second half of next century",
       x = NULL,
       y = NULL) +
  coord_flip() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(-1, 1, by = 0.2)) +
  theme(axis.text = element_text(size = 14),
        legend.position = "bottom")
ggsave(paste0(country_results_dir,
                "/flood_freq_second_half_century", "_", flood_size, climate_future, ".pdf"),
    flood_second_half_stackedbar, 
    width = 20,
    height = 7,
    dpi = 1000)


write.csv(tabulated_floods, 
          file = paste0(dir, "/tabulated_floods_",
                        flood_size, climate_future, ".csv"),
          row.names = FALSE)



```

## Free time dreams

We can dream, right?

```{r dreams, echo=FALSE}

# get (gridded or country-level) population projections and use those to recalculate exposed flood persons. Apparently James is working on this

# use finer resolution flood exposures from Kimura: "decreasing the spatial resolution of the flood-hazard map from 90 to 900m increased the exposure by 51%–94% for different population products"

# try a sensitivity analysis of how results change considering population growth (with or without changes in population distribution)

# find a measure of vulnerability (at country level) and apply to the calculation of flood-related deaths

# use the SDI-specific relative risks (one provided for low SDI countries and the other for high SDI countries). From Hirabayashi paper: "previous studies have demonstrated that vulnerability to flooding has changed over time in association with economic development."

# make 1,000 draws of flood futures (using annual exceedance probability = 1 / return period? or a different formula/concept?) to be able to calculate uncertainty intervals. See comment from Laura below

# reflect the increased hazard (RR) expected at later time points in worse climate scenarios based on IPCC (suggested that warming temperatures are also expected to increase the hazard of a given flooding event independent of flood frequency or numbers of people exposed. Projected increases in direct flood damages are higher by 1.4 to 2 times at 2°C and 2.5 to 3.9 times at 3°C compared to 1.5°C global warming without adaptation).

```
