---
title: "data_exploration"
author: "Susan Campbell"
date: "2024-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data descriptions

### Exposure data

These data were provided by Dr. Yukiko Hirabayashi's team. Key features:

-   Temporal coverage and resolution: 2071-2100 with resolution 0.25 \* 0.25 degrees (

-   Spatial coverage and resolution: globe with high resolution (point data)

-   Measure of flood frequency: Return period (number of years) for 10-, 30- and 100-year flood events (according to historical 1971-2000 data)

-   Measure of flood intensity: none

-   Future scenarios: SSP126 ("sustainability" at 1.0-1.8 C warming by 2100), SSP 245 ("middle of road" at 1.3-2.4 C by 2100), and SSP 585 ("fossil fuels" at 3.3-5.7 C by 2100)

For more information about its origins, see: Hirabayashi, Y., Tanoue, M., Sasaki, O., Zhou, X., & Yamazaki, D. (2021). Global exposure to flooding from the new CMIP6 climate model projections. *Scientific Reports*, *11*(1), 3740.

### Hazard data

Describe!

### Country and region data

Describe!

### Population data

Describe!

## Load libraries and set directory

```{r libs, echo=FALSE}

library(tidyverse)
library(readr)
library(rmapshaper)
library(viridis)
library(cowplot)
library(parallel)
library(RColorBrewer)
library(purrr)
library(forecast)

# specify directories
dir <- "/Users/SusanCampbell/UW_HMS_R/capstone/"
setwd(dir) 

# load functions in my process_input_functions.R
source(paste0(dir, "/processing_functions.R")) 

```

Set values related to return period and time window

```{r set_time, echo=FALSE}

# set values related to scenario
initial_return_period <- 100
climate_future <- "SSP585"
flood_size <- paste0(initial_return_period, "yr")

# set number of draws for bootstrapping
num_draws <- 100

# set seed for reproducibility
set.seed(48)

# set values related to time of interest
time_window <- 100
start_year <- 2024
projection_year <- 2085 # midway between 2071 and 2100
end_year <- start_year + time_window

# locations to save results
country_results_dir <- paste0(dir, "results/", flood_size, climate_future, "/country")

```

## Load data and set params

```{r load_data, echo=FALSE}

# load country shapefiles and region hierarchies ---------------------

# from IHME
# no need to load country_boundaries now - load before doing country-level aggregation
# country_boundaries <- vect("data/GBD2023_analysis_final/GBD2023_analysis_final.shp")

# GBD location hierarchies: Global burden of disease (GBD) data and Tools Guide: Full list of locations, diseases, injuries, and risk factors. Institute for Health Metrics and Evaluation. (2024). https://www.healthdata.org/research-analysis/about-gbd/gbd-data-and-tools-guide
location_hierarchy <- read_csv("data/gbd_location_hierarchy.csv")


# load in population data --------------------------------------------

# from WorldPop, processed and corrected by IHME
# load population data from with 0.0417 * 0.0417 degree resolution (~5km at equator)
pop <- rast(paste0(dir, "data/worldpop_total_1y_2024_00_00.tif"))


# read in fraction of country exposed to floods from generate_floods_exposures.Rmd
fraction_country_exposed <- read.csv(paste0(country_results_dir, 
                                              "/fraction_country_exposed_",
                                              flood_size, climate_future, ".csv"))

```

Set hazard values

```{r set_hazard, echo=FALSE}

# hazard values without considering differenital vulnerability ----------------------
# hazard for all-cause mortality
rr_allcause_est <- 1.04
rr_allcause_lower <- 1.02
rr_allcause_upper <- 1.05
# hazard for morality related to respiratory causes
rr_resp_est <- 1.1
rr_resp_lower <- 1.02
rr_resp_upper <- 1.19
# hazard for morality related to cardiovascular causes
rr_cardio_est <- 1.03
rr_cardio_lower <- 1.02
rr_cardio_upper <- 1.05

# set the proportion of a year that the RR and PAF applies to
paf_proportion_year <- 60 / 365

```

Specify whether to write intermediate results to tif files, whether to generate figures, etc.

```{r read_and_write_settings, echo=FALSE}

# if TRUE, country-level results will be read in from tif and csv files instead of generated anew
read_country_results <- FALSE
# if TRUE, country-level results will be saved as tif and/or csv files
write_country_results <- TRUE
# if TRUE, figures (maps and histograms) of country-level results will be saved as pdf files
generate_country_figures <- TRUE

```

## Country results

Process country info

```{r country_info, echo=FALSE}

# read in and process country boundaries shapefile -----------------------------------------

# # country_boundaries <- st_read("data/GBD2023_analysis_final/GBD2023_analysis_final.shp")
# country_boundaries <- st_read("data/WB_countries_Admin0_10m/WB_countries_Admin0_10m.shp")
country_boundaries <- st_read("data/GBD_countries/lbd_standard_admin_0_simplified.shp")
names(country_boundaries) <- c("code", "name", "loc_id", "geo_id", "geometry")

country_id_names_dict <- country_boundaries %>%
  as.data.frame() %>%
  dplyr::select("code", "name", "loc_id") %>%
  mutate(index = row_number())

# write.csv(country_id_names_dict,
#             file = paste0("country_dict.csv"),
#             row.names = FALSE)


# plot country boundaries to check ---------------------------------------------------------

# plot the country boundaries
ggplot() +
  geom_sf(data = country_boundaries, fill = "white", color = "black") +
  theme_minimal() +
  labs(title = "Country Boundaries",
       x = NULL,
       y = NULL)

```

### Country pop and mortality data

Bring in population and mortality data. For mortality, it's all-cause mortality for all IHME-defined countries for each year through 2100

```{r load_pop_mortality_data, echo=FALSE}

# list of countries in GBD location names that we want to keep
locids_to_keep <- unique(country_id_names_dict$loc_id)


# read in country population forecasts for 2017 through 2100
pop_countries_df <- read.csv(paste0(dir, "data/GBD_pop_forecast_2017_2100.csv"))

pop_countries <- extend_forecasts(pop_countries_df, 
                                  ages = "All Ages", locs = locids_to_keep,
                                  end_year = end_year)

pop_draws <- execute_all_draws(pop_countries, prefix = "pop_draw_")
names(pop_draws)[3:5] <- c("population", "population_lower", "population_upper")


# read in country all-cause mortality forecasts for 2017 through 2100
all_cause_mortality_df <- read.csv(paste0(dir, "data/GBD_mortality_forecast_2017_2100.csv"))

all_cause_mortality <- extend_forecasts(all_cause_mortality_df,
                                        ages = "All Ages", locs = locids_to_keep,
                                        end_year = end_year)

# generate draws for mort dataframe
mort_draws <- execute_all_draws(all_cause_mortality, prefix = "mort_draw_")
names(mort_draws)[3:5] <- c("deaths", "deaths_lower", "deaths_upper")


# combine the draws for population and mortality
combined_draws <- full_join(pop_draws, mort_draws, by = c("loc_id", "year"))

# perform calculations
calculated_draws <- combined_draws %>%
  dplyr:: select(!c("population", "population_lower", "population_upper",
                    "deaths", "deaths_lower", "deaths_upper")) %>%
  pivot_longer(cols = starts_with("pop_draw_"), 
               names_to = "pop_draw_id", values_to = "population") %>%
  pivot_longer(cols = starts_with("mort_draw_"), 
               names_to = "mort_draw_id", values_to = "deaths") %>%
  mutate(draw_id = as.integer(gsub("pop_draw_", "", pop_draw_id))) %>%
  filter(draw_id == as.integer(gsub("mort_draw_", "", mort_draw_id))) %>%
  mutate(mortality_rate = deaths / population) %>%
  dplyr::select(-pop_draw_id, -mort_draw_id)

# correct any negative values to 0
calculated_draws$population[calculated_draws$population < 0] <- 0
calculated_draws$deaths[calculated_draws$deaths < 0] <- 0
calculated_draws$mortality_rate[calculated_draws$mortality_rate < 0] <- 0

# calculate PAF from RR
paf_allcause_est <- calculate_paf_from_rr(rr_allcause_est, 
                                          proportion_year = paf_proportion_year)
paf_allcause_lower <- calculate_paf_from_rr(rr_allcause_lower, 
                                            proportion_year = paf_proportion_year)
paf_allcause_upper <- calculate_paf_from_rr(rr_allcause_upper, 
                                            proportion_year = paf_proportion_year)

# join with country_id_names_dict
forecasted_deaths <- calculated_draws %>%
  left_join(country_id_names_dict, by = "loc_id") %>%
  filter(year >= start_year & year <= end_year) %>%
  rowwise() %>%
  mutate(paf = generate_draw(paf_allcause_est, paf_allcause_lower, paf_allcause_upper, 
                             ndraws = 1)) %>%
  ungroup()

```

### Country-level deaths and death rate

Calculate excess deaths and excess death rates for countries

```{r calculate_deaths_and_rates, echo=FALSE}


if(read_country_results == TRUE) {
  
  country_yearly_findings <- read.csv(file = paste0(country_results_dir, 
                                                     "/country_yearly_findings_",
                                                     flood_size, climate_future, ".csv"))
  
  country_yearly_findings_uncertainty <- read.csv(file = paste0(country_results_dir, 
                                                     "/country_yearly_findings_uncertainty_",
                                                     flood_size, climate_future, ".csv"))
} else {
  
  # # prepare df of 2024 population counts
  # population_2024 <- forecasted_deaths %>%
  #   filter(year == 2024) %>%
  #   group_by(loc_id) %>%
  #   summarize(pop_2024 = mean(population, na.rm = TRUE))
  
  
  # function to assemble country yearly findings from forecasted pop and mort and flood exposures
  assemble_country_findings <- function(exposure_df, forecast_df, pop_2024) {

    findings_df <- exposure_df %>%
      full_join(forecast_df, join_by(name, loc_id, year)) %>%
      #left_join(pop_2024, join_by(loc_id)) %>%
      #mutate(fraction_exposed = number_exposed / pop_2024) %>% 
      mutate(flood_attributable_death_rate = mortality_rate * fraction_exposed * paf,
            flood_attributable_deaths = deaths * fraction_exposed * paf)
      
    return(findings_df)
  }
  
  
  # assemble findings for all countries and all years based on 50th percentile estimates
  country_yearly_findings <- assemble_country_findings(exposure_df = fraction_country_exposed,
                                                       forecast_df = forecasted_deaths,
                                                       pop_2024 = population_2024)
  
  calculate_percentiles <- function(df, death_col = "flood_attributable_deaths", 
                                  deathrate_col = "flood_attributable_death_rate") {
    df_to_join <- df %>%
      dplyr::select(c(loc_id, name, year, exposed, fraction_exposed)) %>%
      unique()
  
    df_summary <- df %>%
      group_by(year, name) %>%
      summarize(
        flood_attributable_deaths_lower = quantile(get(death_col), 0.025, na.rm = TRUE),
        flood_attributable_deaths_upper = quantile(get(death_col), 0.975, na.rm = TRUE),
        flood_attributable_deaths = quantile(get(death_col), 0.5, na.rm = TRUE),
        flood_attributable_death_rate_lower = quantile(get(deathrate_col), 0.025, na.rm = TRUE),
        flood_attributable_death_rate_upper = quantile(get(deathrate_col), 0.975, na.rm = TRUE),
        flood_attributable_death_rate = quantile(get(deathrate_col), 0.5, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      ungroup() %>%
      left_join(df_to_join, join_by(name, year))
    return(df_summary)
  }

  country_yearly_findings_uncertainty <- calculate_percentiles(country_yearly_findings)
  country_yearly_findings <- country_yearly_findings %>%
    group_by(name, year) %>%
    summarize(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) %>%
    dplyr::select(!paf)

}


if(write_country_results == TRUE) {
  
  write.csv(country_yearly_findings, 
            file = paste0(country_results_dir, "/country_yearly_findings_",
                          flood_size, climate_future, ".csv"),
            row.names = FALSE)
  
  write.csv(country_yearly_findings_uncertainty, 
            file = paste0(country_results_dir, "/country_yearly_findings_uncertainty_",
                          flood_size, climate_future, ".csv"), 
            row.names = FALSE)
}

```

Plots to check and visualize flood-attributable deaths

```{r plot_deaths, echo=FALSE}

# plots to check --------------------------------------------------------------------

country_yearly_with_boundaries <- country_boundaries %>%
  left_join(country_yearly_findings, join_by(name, loc_id, code))

early_year <- 3
late_year <- 97

map_deaths_early <- plot_excess_deaths_map(country_yearly_with_boundaries, 
                                           year = early_year,
                                           title = paste("Flood-attributable all-cause deaths,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", early_year, ")"),
                                           fixed_binned_scales = FALSE)
map_deaths_early_comparable <- plot_excess_deaths_map(country_yearly_with_boundaries, 
                                           year = early_year,
                                           title = paste("Flood-attributable all-cause deaths,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", early_year, ")"),
                                           fixed_binned_scales = TRUE)

map_deaths_late <- plot_excess_deaths_map(country_yearly_with_boundaries, 
                                           year = late_year,
                                           title = paste("Flood-attributable all-cause deaths,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", late_year, ")"),
                                           fixed_binned_scales = FALSE)
map_deaths_late_comparable <- plot_excess_deaths_map(country_yearly_with_boundaries, 
                                           year = late_year,
                                           title = paste("Flood-attributable all-cause deaths,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", late_year, ")"),
                                           fixed_binned_scales = TRUE)


if(generate_country_figures == TRUE) {
  
  ggsave(paste0(country_results_dir,
                "/map_deaths_year", early_year, "_", flood_size, climate_future, ".pdf"),
    map_deaths_early, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(country_results_dir,
                "/map_deaths_year", late_year, "_", flood_size, climate_future, ".pdf"),
    map_deaths_late, 
    width = 9,
    height = 4,
    dpi = 1000)
  
    ggsave(paste0(country_results_dir,
                "/map_deaths_year", early_year, "_", flood_size, climate_future, "_comparable.pdf"),
    map_deaths_early_comparable, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(country_results_dir,
                "/map_deaths_year", late_year, "_", flood_size, climate_future, "_comparable.pdf"),
    map_deaths_late_comparable, 
    width = 9,
    height = 4,
    dpi = 1000)
  
}



deaths_over_time_data <- country_yearly_findings %>%
  group_by(name) %>%
  summarize(
    total_deaths = sum(flood_attributable_deaths, na.rm = TRUE)) %>%
  right_join(country_boundaries, join_by(name,))


deaths_over_time <- plot_deaths_over_time_map(deaths_over_time_data,
                                                  title = paste("Sum of people exposed to", 
                                                                flood_size,"floods, over next",
                                                                time_window, "years, in",
                                                                climate_future),
                                                  fixed_binned_scales = FALSE)

deaths_over_time_comparable <- plot_deaths_over_time_map(deaths_over_time_data,
                                                  title = paste("Sum of people exposed to", 
                                                                flood_size,"floods, over next",
                                                                time_window, "years, in",
                                                                climate_future),
                                                  fixed_binned_scales = TRUE)

if(generate_country_figures == TRUE) {
  
  ggsave(paste0(country_results_dir,
                "/map_deaths_100_years", "_", flood_size, climate_future, ".pdf"),
    deaths_over_time, 
    width = 9,
    height = 4,
    dpi = 1000)
  
    ggsave(paste0(country_results_dir,
                "/map_deaths_100_years", "_", flood_size, climate_future, "_comparable.pdf"),
    deaths_over_time_comparable, 
    width = 9,
    height = 4,
    dpi = 1000)
}

```

Plots to check and visualize flood-attributable death rate

```{r plot_death_rates, echo=FALSE}

# plots to check --------------------------------------------------------------------

country_yearly_with_boundaries <- country_boundaries %>%
  left_join(country_yearly_findings, join_by(name, loc_id, code))

early_year <- 3
late_year <- 97

map_death_rate_early <- plot_excess_death_rate_map(country_yearly_with_boundaries, 
                                           year = early_year,
                                           title = paste("Flood-attributable all-cause death rate,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", early_year, ")"),
                                           fixed_binned_scales = FALSE)
map_death_rate_early_comparable <- plot_excess_death_rate_map(country_yearly_with_boundaries, 
                                           year = early_year,
                                           title = paste("Flood-attributable all-cause death rate,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", early_year, ")"),
                                           fixed_binned_scales = TRUE)

map_death_rate_late <- plot_excess_death_rate_map(country_yearly_with_boundaries, 
                                           year = late_year,
                                           title = paste("Flood-attributable all-cause death rate,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", late_year, ")"),
                                           fixed_binned_scales = FALSE)
map_death_rate_late_comparable <- plot_excess_death_rate_map(country_yearly_with_boundaries, 
                                           year = late_year,
                                           title = paste("Flood-attributable all-cause death rate,",
                                                         flood_size, "floods in", climate_future, 
                                                         "(year", late_year, ")"),
                                           fixed_binned_scales = TRUE)


if(generate_country_figures == TRUE) {
  
  ggsave(paste0(country_results_dir,
                "/map_deathrate_year", early_year, "_", flood_size, climate_future, ".pdf"),
    map_death_rate_early, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(country_results_dir,
                "/map_deathrate_year", late_year, "_", flood_size, climate_future, ".pdf"),
    map_death_rate_late, 
    width = 9,
    height = 4,
    dpi = 1000)
  
    ggsave(paste0(country_results_dir,
                "/map_deathrate_year", early_year, "_", flood_size, climate_future, "_comparable.pdf"),
    map_death_rate_early_comparable, 
    width = 9,
    height = 4,
    dpi = 1000)
  
  ggsave(paste0(country_results_dir,
                "/map_deathrate_year", late_year, "_", flood_size, climate_future, "_comparable.pdf"),
    map_death_rate_late_comparable, 
    width = 9,
    height = 4,
    dpi = 1000)
  
}

```

## Analyses and visualizations

```{r scenario_summary_plots, echo=FALSE}

# stitch together maps ---------------------------------------------------------------------

# create a list of the plots to include
plots <- list(map_deaths_early, map_deaths_late, deaths_over_time, 
              map_death_rate_early, map_death_rate_late)

scenario_summary_plot <- plot_grid(plotlist = plots, ncol = 3, nrow = 2)

# save as pdf
save_plot(filename = paste0(country_results_dir, "/scenario_summary_", 
                            flood_size, "_", climate_future,".pdf"), 
          plot = scenario_summary_plot, 
          dpi = 2400,
          base_height = 13, base_width = 30)

# with comparable, fixed binned scales ------------------------------------------------------

# create a list of the plots to include
plots_comp <- list(map_deaths_early_comparable, map_deaths_late_comparable,
                   deaths_over_time_comparable,
                   map_death_rate_early_comparable, map_death_rate_late_comparable)

scenario_summary_plot_comp <- plot_grid(plotlist = plots_comp, ncol = 3, nrow = 2)

# save as pdf
save_plot(filename = paste0(country_results_dir, "/scenario_summary_comparable_", 
                            flood_size, "_", climate_future,".pdf"), 
          plot = scenario_summary_plot_comp, 
          dpi = 2400,
          base_height = 13, base_width = 30)

```
